---
title: "Data Cleaning"

format:
  html:
    df-print: paged
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true

editor: 
  markdown: 
    wrap: 72
---

# Load libraries

```{r}
#| label: setup
#| warning: false
#| message: false
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
source('dependencies.R')

```

# Read in Data Sets

## State Item Analysis Report
```{r}
HSPhy_2022_StateItemDF<-read_state_item("data/2022_Physics_District_NextGenMCASItem.xlsx", 2022, "PHY" )

HSPhy_2022_StateItemDF

HSPhy_2023_StateItemDF<-read_state_item("data/2023_Physics_District_NextGenMCASItem.xlsx", 2023, "PHY" )

HSPhy_2023_StateItemDF



```

## School Item Analysis Reports

```{r}
HSPhy_2022_SchoolItemDF<-read_school_item("data/2022_Physics_NextGenMCASItem.xlsx", 2022, "PHY")
HSPhy_2022_SchoolItemDF

HSPhy_2022_SchoolItemDF<-HSPhy_2022_SchoolItemDF%>%
  left_join(HSPhy_2022_StateItemDF, by= c('Year'='Year', 'Subject'='Subject', 'ITEM' = 'ITEM'))%>%
  mutate(`School-State Diff` = `School%`- `State%`)
HSPhy_2022_SchoolItemDF


HSPhy_2023_SchoolItemDF<-read_school_item("data/2023_Physics_NextGenMCASItem.xlsx", 2023, "PHY")
HSPhy_2023_SchoolItemDF

HSPhy_2023_SchoolItemDF<-HSPhy_2023_SchoolItemDF%>%
  left_join(HSPhy_2023_StateItemDF, by= c('Year'='Year', 'Subject'='Subject', 'ITEM' = 'ITEM'))%>%
  mutate(`School-State Diff` = `School%`- `State%`)

HSPhy_2023_SchoolItemDF

tail(HSPhy_2023_SchoolItemDF)

tail(HSPhy_2022_SchoolItemDF)

HSPhy_NextGen_SchoolItemDF <- rbind(HSPhy_2022_SchoolItemDF, HSPhy_2023_SchoolItemDF)

HSPhy_NextGen_SchoolItemDF






```

## Summary State NextGen Achievement Reports

```{r}



HSPhy_2022_StateAchievementDF<-read_state_achievement("data/2022_HSSci_NextGenMCAS.xlsx", 2022, "PHY" )
HSPhy_2022_StateAchievementDF

HSPhy_2023_StateAchievementDF<-read_state_achievement("data/2023_HSSci_NextGenMCAS.xlsx", 2023, "PHY" )
HSPhy_2023_StateAchievementDF

HSPhy_NextGen_StateAchievementDF<-rbind(HSPhy_2022_StateAchievementDF, HSPhy_2023_StateAchievementDF)
HSPhy_NextGen_StateAchievementDF
length(unique(HSPhy_NextGen_StateAchievementDF$`School Code`))

HSPhy_NextGen_StateSummaryDF<- HSPhy_NextGen_StateAchievementDF%>%
  summarise(`State Tested` = sum(`State Tested`),
            `E#State` = sum(`E#State`),
            `M#State` = sum(`M#State`),
            `PM#State` = sum(`PM#State`),
            `NM#State` = sum(`NM#State`),
            `Score Total` = sum(`State Avg. Scaled Score`)
            )%>%
  mutate(`E%State` = round(`E#State`/`State Tested`*100))%>%
  mutate(`M%State` = round(`M#State`/`State Tested`*100))%>%
  mutate(`PM%State` = round(`PM#State`/`State Tested`*100))%>%
  mutate(`NM%State` = round(`NM#State`/`State Tested`*100))%>%
  #mutate(`State Avg. Score` = round(`Score Total`/`State Tested`))
  mutate(`Subject` = "PHY")%>%
  select(`Subject`, `E%State`, `M%State`, `PM%State`, `NM%State`)

HSPhy_NextGen_StateSummaryDF

HSPhy_NextGen_SchoolSummaryDF<- HSPhy_NextGen_StateAchievementDF%>%
    group_by(`School Code`, `School Name`)%>%
    summarise(`Tested Students` = sum(`Tested Students`),
            `E#` = sum(`E#`),
            `M#` = sum(`M#`),
            `PM#` = sum(`PM#`),
            `NM#` = sum(`NM#`)
            )%>%
    mutate(`E%` = round(`E#`/`Tested Students`*100))%>%
    mutate(`M%` = round(`M#`/`Tested Students`*100))%>%
    mutate(`PM%` = round(`PM#`/`Tested Students`*100))%>%
    mutate(`NM%` = round(`NM#`/`Tested Students`*100))%>%
    mutate(`Subject` = "PHY")%>%
    select(`Subject`, `School Name`, `School Code`, `Tested Students`, `E%`, `M%`, `PM%`, `NM%`)

HSPhy_NextGen_SchoolSummaryDF

HSPhy_NextGen_AchSumDF<-left_join(HSPhy_NextGen_SchoolSummaryDF, HSPhy_NextGen_StateSummaryDF, by = "Subject")%>%
  mutate(`EDiff` = `E%` - `E%State`)%>%
  mutate(`MDiff` = `M%` - `M%State`)%>%
  mutate(`PMDiff` = `PM%` - `PM%State`)%>%
  mutate(`NMDiff` = `NM%` - `NM%State`)
  

HSPhy_NextGen_AchSumDF





```

## IT301 Reports

```{r}

SG9_standardXWalk<-read_excel("data/NextGenMCASItemxWalk.xlsx", sheet = "HS_Phys_StandardxWalk")

IT301_test<-read_excel("data/2023_Physics_IT301 MCAS District and School Test Item Analysis Summary.xlsx", skip = 14)
IT301_test


IT301_2022<-read_IT301("data/2022_Physics_IT301 MCAS District and School Test Item Analysis Summary.xlsx", 2022, "PHY" )
IT301_2022

IT301_2023<-read_IT301("data/2023_Physics_IT301 MCAS District and School Test Item Analysis Summary.xlsx", 2023, "PHY" )
IT301_2023

NextGenIT301<- rbind(IT301_2022, IT301_2023)

HSPhy_NextGenIT301<-NextGenIT301%>%
  left_join(SG9_standardXWalk, by = "Standard")
HSPhy_NextGenIT301

HSPhy_NextGen_PCSum<-HSPhy_NextGenIT301%>%
 group_by(`Practice Category`)%>%
  summarise(`Practice Cat Pts` = sum(`Item Possible Points`),
          `State Practice Cat Pts` = sum(`Item Possible Points`*`State%` ))%>%
    mutate(`State Practice Cat%` = round(100*`State Practice Cat Pts`/ `Practice Cat Pts`))%>%
  select(`Practice Category`, `State Practice Cat%`)%>%
  pivot_wider(names_from = `Practice Category`, values_from = `State Practice Cat%`)%>%
  mutate(`MD%State` =`B. Mathematics and Data`) %>%
  mutate(`IQ%State` = `A. Investigations and Questioning`)%>%
  mutate(`ERM%State` =`C. Evidence, Reasoning, and Modeling`)%>%
  mutate(`None%State` = `None`)%>%
  mutate(`Subject` = "PHY")%>%
  select(`Subject`, `IQ%State`, `MD%State`, `ERM%State`, `None%State`)

HSPhy_NextGen_PCSum


HSPhy_NextGen_RCSum<-HSPhy_NextGenIT301%>%
 group_by(`Reporting Category`)%>%
  summarise(`Report Cat Pts` = sum(`Item Possible Points`),
          `State Report Cat Pts` = sum(`Item Possible Points`*`State%` ))%>%
    mutate(`State Report Cat%` = round(100*`State Report Cat Pts`/ `Report Cat Pts`))%>%
  select(`Reporting Category`, `State Report Cat%`)%>%
  pivot_wider(names_from = `Reporting Category`, values_from = `State Report Cat%`)%>%
   mutate(`EN%State` =`EN`) %>%
   mutate(`MF%State` = `MF`)%>%
   mutate(`WA%State` =`WA`)%>%
   mutate(`Subject` = "PHY")%>%
   select(`Subject`, `EN%State`, `MF%State`, `WA%State`)

HSPhy_NextGen_RCSum

HSPhy_NextGen_CatSum <- left_join(HSPhy_NextGen_RCSum, HSPhy_NextGen_PCSum, by = "Subject")
HSPhy_NextGen_CatSum

HSPhy_NextGen_SumDF<-left_join(HSPhy_NextGen_AchSumDF, HSPhy_NextGen_CatSum, by = "Subject")

HSPhy_NextGen_SumDF

```

## Join IT301 Summary reports to SchoolItem performance Reports

```{r}



HSPhy_NextGen_SchoolIT301DF <- left_join(HSPhy_NextGen_SchoolItemDF, HSPhy_NextGenIT301, by = c("Year" = "Year", "Subject" = "Subject", "ITEM" = "ITEM"))

HSPhy_NextGen_SchoolIT301DF
view(HSPhy_NextGen_SchoolIT301DF)
length(unique(HSPhy_NextGen_SchoolIT301DF$`School Name`))

```

## Compute Key Summary Stats for IT301 Reports

Data of 112 schools performance on Next Generation HS Introductory Physics exam in Spring of 2022 and Spring of 2023. 87 of those schools tested students in both years and 25 of those schools only tested students in 1 of the 2 testing years.
```{r}
## Compute % Earned by Practice Category
## Compute State % Earned by Practice Category
## Compute SD by Practice Category
## Compute State % Earned by Reporting Category
## Compute SD by Reporting Category

Practice_Cat_Sum <- HSPhy_NextGen_SchoolIT301DF%>%
  group_by(`Practice Category`)%>%
  summarise(`PC Mean Diff` = mean(`School-State Diff`),
            `PC Med Diff` = median(`School-State Diff`),
            `PC SD Diff` = sd(`School-State Diff`))

Practice_Cat_Sum

Practice_Cat_School_Sum <- HSPhy_NextGen_SchoolIT301DF%>%
  group_by(`School Name`,  `School Code`, `Practice Category`)%>%
  summarise(`PC Mean Diff` = mean(`School-State Diff`),
            `PC Med Diff` = median(`School-State Diff`),
            `PC SD Diff` = sd(`School-State Diff`))

Practice_Cat_School_Sum


Reporting_Cat_Sum <- HSPhy_NextGen_SchoolIT301DF%>%
  group_by(`Reporting Category`)%>%
  summarise(`RC Mean Diff` = mean(`School-State Diff`),
            `RC Med Diff` = median(`School-State Diff`),
            `RC SD Diff` = sd(`School-State Diff`))

Reporting_Cat_Sum

Reporting_Cat_School_Sum <- HSPhy_NextGen_SchoolIT301DF%>%
  group_by(`School Name`,  `School Code`, `Reporting Category`)%>%
  summarise(`RC Mean Diff` = mean(`School-State Diff`),
            `RC Med Diff` = median(`School-State Diff`),
            `RC SD Diff` = sd(`School-State Diff`))


Reporting_Cat_School_Sum





```


## State E below average with Practice Category

```{r}
# State E below average
HSPhy_NextGen_StateAchievementDF

LowEStateDF<- HSPhy_NextGen_StateAchievementDF%>%
  filter(`Performance Level` == "E")%>%
  filter(`Performance Diff` < 0)%>%
  mutate(`ELow` = TRUE)

LowEStateDF
#view(LowEStateDF)

LowEPractice<-LowEStateDF%>%
  select(`Year`, `Subject`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, `State Avg. Scaled Score`)%>%
  left_join(Practice_Cat_School_Sum, by = c( "School Code" = "School Code"))%>%
  select(`Year`, `School Name`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, 
         `State Avg. Scaled Score`, `Practice Category`, `PC Mean Diff`, `PC Med Diff`, `PC SD Diff`)

LowEPractice%>%
  group_by(`Practice Category`)%>%
  summarise(`Low E PC Mean Diff` = mean(`PC Mean Diff`, na.rm = TRUE),
            `Low E PC Med Diff` = median(`PC Med Diff`, na.rm = TRUE),
            `Low E PC SD Diff` = sd(`PC SD Diff`, na.rm = TRUE)
            )
```


## State E above average with Practice Category

```{r}
HSPhy_NextGen_StateAchievementDF

HighEStateDF<- HSPhy_NextGen_StateAchievementDF%>%
  filter(`Performance Level` == "E")%>%
  filter(`Performance Diff` > 0)%>%
  mutate(`ELow` = FALSE)

HighEStateDF
#view(HighEStateDF)

HighEPractice<-HighEStateDF%>%
  select(`Year`, `Subject`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, `State Avg. Scaled Score`)%>%
  left_join(Practice_Cat_School_Sum, by = c( "School Code" = "School Code"))%>%
  select(`Year`, `School Name`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, 
         `State Avg. Scaled Score`, `Practice Category`, `PC Mean Diff`, `PC Med Diff`, `PC SD Diff`)

HighEPractice%>%
  group_by(`Practice Category`)%>%
  summarise(`High E PC Mean Diff` = mean(`PC Mean Diff`, na.rm = TRUE),
            `High E PC Med Diff` = median(`PC Med Diff`, na.rm = TRUE),
            `High E PC SD Diff` = sd(`PC SD Diff`, na.rm = TRUE)
            )
```


## State E and M above average

```{r}
HSPhy_NextGen_StateAchievementDF

HighMEStateDF<- HSPhy_NextGen_StateAchievementDF%>%
  filter(`Performance Level` == "E" | `Performance Level` == "M")%>%
  filter(`Performance Diff` > 0)%>%
  mutate(`MELow` = FALSE)

HighMEStateDF
#view(HighEStateDF)

HighMEPractice<-HighMEStateDF%>%
  select(`Year`, `Subject`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, `State Avg. Scaled Score`)%>%
  left_join(Practice_Cat_School_Sum, by = c( "School Code" = "School Code"))%>%
  select(`Year`, `School Name`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, 
         `State Avg. Scaled Score`, `Practice Category`, `PC Mean Diff`, `PC Med Diff`, `PC SD Diff`)

HighMEPractice%>%
  group_by(`Practice Category`)%>%
  summarise(`High ME PC Mean Diff` = mean(`PC Mean Diff`, na.rm = TRUE),
            `High ME PC Med Diff` = median(`PC Med Diff`, na.rm = TRUE),
            `High ME PC SD Diff` = sd(`PC SD Diff`, na.rm = TRUE)
            )
```


## State E or M below average


```{r}
HSPhy_NextGen_StateAchievementDF

LowMEStateDF<- HSPhy_NextGen_StateAchievementDF%>%
  filter(`Performance Level` == "E" | `Performance Level` == "M")%>%
  filter(`Performance Diff` < 0)%>%
  mutate(`MELow` = FALSE)

LowMEStateDF
#view(LowEStateDF)

LowMEPractice<-LowMEStateDF%>%
  select(`Year`, `Subject`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, `State Avg. Scaled Score`)%>%
  left_join(Practice_Cat_School_Sum, by = c( "School Code" = "School Code"))%>%
  select(`Year`, `School Name`, `School Code`, `Tested Students`, `Performance Level`, `Performance Count`, `Performance%`, `State Performance%`, `Performance Diff`, `Avg. Scaled Score`, 
         `State Avg. Scaled Score`, `Practice Category`, `PC Mean Diff`, `PC Med Diff`, `PC SD Diff`)

LowMEPractice%>%
  group_by(`Practice Category`)%>%
  summarise(`Low ME PC Mean Diff` = mean(`PC Mean Diff`, na.rm = TRUE),
            `Low ME PC Med Diff` = median(`PC Med Diff`, na.rm = TRUE),
            `Low ME PC SD Diff` = sd(`PC SD Diff`, na.rm = TRUE)
            )
```


## State E below, M above

```{r}
HSPhy_NextGen_StateAchievementDF

LowEHighMStateDF<-HSPhy_NextGen_StateAchievementDF%>%
  mutate(`MHigh` = case_when(`Performance Level` == "M" & `Performance Diff` >= 0 ~ TRUE,
                             `Performance Level` == "M" & `Performance Diff` < 0 ~ FALSE))%>%
  mutate(`ELow` = case_when(`Performance Level` == "E" & `Performance Diff` < 0 ~ TRUE,
                            `Performance Level` == "E" & `Performance Diff` >= 0 ~ FALSE))%>%
  mutate(`EMOpp` = case_when(`MHigh` == TRUE && `ELow` == TRUE ~ TRUE))#%>%
  #filter(`EMOpp` == TRUE)#%>%
  #filter(`MHigh` == TRUE)
  

view(LowEHighMStateDF)
```