---
title: "DACSS603Final"
author: "Theresa Szczepanski"
desription: "MCAS G9 Science Analysis"
date: "10/22/2023"

format:
  html:
    df-print: paged
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
    
bibliography: references.bib

editor: 
  markdown: 
    wrap: 72
---



```{r}
#| label: setup
#| warning: false
#| message: false

source('dependencies.R')
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


# Research Questions

[@knuth84]

Massachusetts Comprehensive Assessment System (MCAS) tests were
introduced as part of the Massachusetts Education Reform Act in 1993
with the goal of providing all students with the skills and knowledge to
thrive in a "complex and changing society" [@Boats20].
The MCAS tests are a significant tool for educational equity. Scores on
the Grade 10 Math MCAS test "predict longer-term educational attainments
and labor market success, above and beyond typical markers of student
advantage. For example, among demographically similar students who
attended the same high school and have the same level of ultimate
educational attainment, those with higher MCAS mathematics scores go on
to have much higher average earnings than those with lower scores."
[@Boats20].

The Massachusetts Education Reform Act in 1993 was passed in the context of a 
national movement toward education reform throughout the United States. 
As early as 1989 there were calls to establish national curriculum standards 
as a way to improve student college and career readiness skills and close poverty 
gaps (Greer, 2018, 106). With the introduction of the new Common Core standards came the demand for appropriate curricular materials and teaching practices. Four out of five math 
teachers in five studied states, including Massachusetts, reported changing 
more than half of their instructional materials in response to the 
Common Core (Kane et al., 2016). Most United States public schools use commercially produced instructional materials with an estimated 70 to 98 percent of teachers using textbooks at least weekly (Brown Center on Educational Policy at Brookings et al., 2012). Research indicates that the choice of instructional materials can have an impact “as large as or larger than the impact of teacher quality” (Brown Center on Educational Policy at Brookings et al., 2012).

Massachusetts, along with Arkansas, Delaware, Kentucky, Louisiana, Maryland, Mississippi, Nebraska, New Mexico, Ohio, Rhode Island, Tennessee and Texas belongs to the Council of Chief State School Officers' (CCSO), [High Quality Instructional Materials and Professional Development network](https://learning.ccsso.org/high-quality-instructional-materials) with the stated goal of 
ensuring that "Every student everyday is engaged in meaningful, affirming, grade-level instruction, by ensuring that every teacher has access to high-quality, standards aligned instructional materials and receives relevant professional development to support their use of these materials. [Case Study](https://753a0706.flowpaper.com/CCSSOIMPDCaseStudyOVERVIEWFINAL/#page=3)

In DESE's 

When considering Next Generation MCAS High School Introductory Physics student performance data, we hope to address the following broad questions:

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
```
::: blue

-   Is there a relationship between differences in student performance across Science Practice 
    Categories and student acheivement?
    
-   How can trends in student performance be used to identify Introductory Physics 
    content areas in need of curricular adjustments?



:::


In this report, I will analyze the Spring 2022 MCAS Results for students
completing the High School Introductory Physics MCAS at [Rising Tide
Charter Public School](https://risingtide.org/).

The `MCAS_2022` data frame contains performance results from 495
students from Rising Tide on the Spring 2022 [Massachusetts
Comprehensive Assessment System
(MCAS)](https://www.doe.mass.edu/mcas/default.html) tests.

For each student, there are values reported for 256 different variables
which consist of information from four broad categories

-   *Demographic characteristics* of the students themselves (e.g.,
    race, gender, date of birth, town, grade level, years in school,
    years in Massachusetts, and low income, title1, IEP, 504, and EL
    status ).

-   *Key assessment features* including subject, test format, and
    accommodations provided

-   *Performance metrics*: This includes a student's score on individual
    item strands, e.g.,`sitem1`-`sitem42`.

See the `MCAS_2022` data frame summary and **codebook** in the
**appendix** for further details.

The second data set, `SG9_Item`, is $42 \times 9$ and consists of 9
variables with information pertaining to the 42 questions on the 2022
[HS Introductory Physics Item
Report](https://profiles.doe.mass.edu/mcas/mcasitems2.aspx?grade=HS&subjectcode=PHY&linkid=23&orgcode=04830000&fycode=2022&orgtypecode=5&).
The variables can be broken down into 2 categories:

Details about the content of a given test item:

This includes the content `Reporting Category` (MF (motion and forces)
WA (waves), and EN (energy)), the `Standard` from the [2016 STE
Massachusetts Curriculum
Framework](https://www.doe.mass.edu/frameworks/scitech/2016-04.pdf), the
`Item Description` providing the details of what specifically was asked
of students, and the points available for a given question,
`item Possible Points`.

Summary Performance Metrics:

-   For each item, the state reports the percentage of points earned by
    students at Rising Tide, `RT Percent Points`, the percentage of
    available points earned by students in the state,
    `State Percent Points`, and the difference between the percentage of
    points earned by Rising Tide students and the percentage of points
    earned by students in the state, `RT-State Diff`.

-   Lastly, `SG9_CU306Dis` and `SG9_CU306NonDis` are $3 \times 5$
    dataframes consisting of summary performance data by
    `Reporting Category` for students with disabilities and without
    disabilities; most importantly including `RT Percent Points` and
    `State Percent Points`by disability status.


# Hypothesis

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
```
::: blue

-   A school's percent of student's exceeding expectations on the Introductory Physics MCAS 
is negatively associated with variance in Mathematics and Data items.

- School's mean performance diffs are not the same across science practice categories.
    


:::



# Descriptive Statistics

```{r}
# HSPhy_NextGen_PerfDF
# HSPhy_NextGen_SchoolIT301DF

```

# Visualization


# References
